\textbf{Discussion On the Out-of-Sync TDC Module from the HMS Chambers}

### 2018-08-29 Spring_2018_Run/OutOfSync_TDC/tdc_sync_report.tex
finished document on bad TDC

\title{On the Out-of-Sync TDC Module on the HMS Chambers}
\author{Carlos Yero}
\date{August 28 2018}

@@

%*****************************************************************************************************************************************************
During the Fall-2017/Spring-2018 Hall C Commissioning Run Period, it was found that a particular TDC module in the HMS
Drift Chambers Crate (ROC03) was shifting the tdc times of the corresponding groups of wires being fed into it. At that time,
this did not seem to be a serious concern, as this could be fixed in software. There also seem to be no issues with the
residuals per wire after calibration. After the end of the Spring Run, I decided to look at the residuals per wire for a defocused
run in the HMS, and found certaing groups of wire residuals to be significantly smeared compared to adjacent groups of wires.
These were the same group of wires that exhibited the shift in time, which pointed to the same TDC module in ROC03, Slot 2.
In this document I discuss the following: \\
\\
1) the observations made that led to the determination of an out-of-sync TDC module \\
2) what was done to fix this problem in future experiments \\
3) how to minimize the impact on wire residuals for the Spring 2018 Run

\section{Observed Correlations on the Out-of-Sync TDC Module}  
\begin{figure}[h!]
  \centering
  \includegraphics[width=7.0in, height=4.0in]{shifted_drifttimes.png}
  \caption{The Raw TDC vs. Wire Number Correlations for all 12 planes. An offset of $\sim$ 1000 Channels ($\sim$ 100 ns) is observed
    in wire groups of four distinct planes.}
  \label{fig:shift_dtimes}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=7.0in, height=4.0in]{bad_wireresidual.png}
  \caption{Correlation of Residuals vs. Wire Number for all 12 planes. Four wire groups stand out from the rest, with a noticeable smearing.
  The adjacent planes are affected by this smearing as well, even though they are NOT part of the problematic TDC.}
  \label{fig:wire_residual}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=6.0in, height=4.0in]{hDC_RefTime_correlation.png}
  \caption{Correlation between the HMS DC reference time and the raw TDC times. The DC 1, V'-Plane raw TDC vs. Wire Number is shown. Adjacent wirgroups which
  are NOT in TDC Slot 2 are compared to the TDC Slot 2 wire group before and after applying the reference time correction to the TDC time.}
  \label{fig:wire_residual}
\end{figure}

@@

\begin{subfigure}{.5\textwidth}
\hspace*{-5cm}                                                           
  \centering
  \includegraphics[width=2.1\linewidth]{2d_residuals_flagON.png}
  \caption{Residuals vs. Wire Number for all Chamber planes, with sigma-per-wire flag \textit{ON}.}
  \label{fig:2d_res_flagON}
\end{subfigure}
\caption{Residuals vs. Wire Number for all Chamber planes. The \textcolor{red}{RED BOXES} highlight the group of wires that were affected, but had nothing to do with the
  out-of-sync TDCs wire groups.}
\label{fig:wire_residuals_flag}
\end{figure}\\
In subfigure (b), when a sigma of 0.06 was applied to the bad groups (Planes 1u1, 1v2, 2v1, 2u2), the adjacent planes (highlighted in \textcolor{red}{RED})
wire residuals improved significantly. This is due the fact that less weight (larger sigma($\sigma$)) was given to the affected group of wires. \\
\newpage
The plane residuals, however, did not have much improvement after increasing the weight factor. At most, there was a 25 $\mu$m improvevent in the Standard Deviation.
See Figure \ref{fig:plane_residuals}
\begin{figure}[ht!]
  \hspace*{-1.5cm}                                                           
%  \vspace*{10.5cm}                                                           
  \centering
  \includegraphics[width=7.0in, height=4.0in]{residuals_compare.pdf}
  \caption{Plane residuals with the sigma-per-wire flag ON (\textcolor{red}{RED}) and OFF (\textcolor{navyblue}{BLUE})}
  \label{fig:plane_residuals}
\end{figure}
\newline
\noindent It is importnant to note that the sigma-per-wire flag and the parameter file are NOT included in \texttt{hallc$\_$replay}, since this is a very specific scenario that
ONLY applies to the Winter-2017/Spring-2018 Run Period. If users who took data during that time, and would like to try adding this flag and the parameter file,
the instructions are as follows:
\begin{itemize}
\item  Create the parameter file in path: \texttt{/PARAM/HMS/DC/hdc$\_$sigma$\_$per$\_$wire.param} \par
  \begin{minipage}{\linewidth}
    \centering
    \includegraphics[width=15cm]{param.png}
    \captionof{figure}{Snippet of sigma-per-wire parameter file.}
    \label{fig:param}
  \end{minipage}
\item  Include the parameter file in path: \texttt{/DBASE/HMS/general.param} \par
  \begin{minipage}{\linewidth}                                                          
    \centering
    \includegraphics[width=10cm]{gen_parm.png}
    \captionof{figure}{Snippet of general.param file.}
    \label{fig:gen_param}
  \end{minipage}
\end{itemize}
\newpage
\onecolumn
\bibliography{template}
\bibliographystyle{acm}

>>>>

\title{\textbf{Discussion On the Out-of-Sync TDC Module from the HMS Chambers}}
\author{Carlos Yero}
\date{August 29 2018}

@@

%*****************************************************************************************************************************************************
\noindent During the Fall-2017/Spring-2018 Hall C Commissioning Run Period, it was found that a particular TDC module in the HMS
Drift Chambers Crate (ROC03) was shifting the tdc times of the corresponding groups of wires being fed into it. At that time,
this did not seem to be a serious concern, as this could be fixed in software. There also seem to be no issues with the
residuals per wire after calibration. After the end of the Spring Run, I decided to look at the residuals per wire for a defocused
run in the HMS (run 1267), and found certain groups of wire residuals to be significantly smeared compared to adjacent groups of wires (smeared
residuals were being overshadowed by another problem fixed by Mark Jones during Spring 2018 SIDIS, concerning the \texttt{hodoscope starttime}).
These were the same group of wires that exhibited the shift in time, which pointed to the same TDC module in ROC03, Slot 2.
In this document I discuss the following: \\
\\
1) the observations made that led to the determination of an out-of-sync TDC module \\
2) what was done to fix this problem in future experiments \\
3) how to minimize the impact on wire residuals for the Spring 2018 Run

\section{Observed Correlations on the Out-of-Sync TDC Module}  
\begin{figure}[h!]
  \centering
  \includegraphics[width=6.5in, height=3.5in]{shifted_drifttimes.png}
  \caption{The Raw TDC vs. Wire Number Correlations for all 12 planes. An offset of $\sim$ 1000 Channels ($\sim$ 100 ns) is observed
    in wire groups of four distinct planes.}
  \label{fig:shift_dtimes}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=7.0in, height=4.0in]{bad_wireresidual.png}
  \caption{Correlation of Residuals vs. Wire Number for all 12 planes. Four wire groups stand out from the rest, with a noticeable smearing.
  The adjacent planes are affected by this smearing as well, even though they are NOT part of the problematic TDC.}
  \label{fig:wire_residual}
\end{figure}
\begin{figure}[h!]
  \centering
  \includegraphics[width=6.0in, height=4.0in]{hDC_RefTime_correlation.png}
  \caption{Correlation between the HMS DC reference time and the raw TDC times. The DC 1, V'-Plane raw TDC vs. Wire Number is shown. Adjacent wirgroups which
  are NOT in TDC Slot 2 are compared to the TDC Slot 2 wire group before and after applying the reference time correction to the TDC time.}
  \label{fig:wire_residual}
\end{figure}

@@

\begin{subfigure}{.5\textwidth}
\hspace*{-5cm}                                                           
  \centering
  \includegraphics[width=2.1\linewidth]{2d_residuals_flagON.png}
  \caption{Residuals vs. Wire Number for all Chamber planes, with sigma-per-wire flag \textit{ON}.}
  \label{fig:2d_res_flagON}
\end{subfigure}
\caption{Residuals vs. Wire Number for all Chamber planes. The \textcolor{red}{RED BOXES} highlight the group of wires that were affected, but had nothing to do with the
  out-of-sync TDCs wire groups.}
\label{fig:wire_residuals_flag}
\end{figure}\\
In subfigure (b), when a sigma of 0.06 was applied to the bad groups (Planes 1u1, 1v2, 2v1, 2u2), the adjacent planes (highlighted in \textcolor{red}{RED})
wire residuals improved significantly. This is due the fact that less weight (larger sigma($\sigma$)) was given to the affected group of wires. \\
\newpage
The plane residuals, however, did not have much improvement after increasing the weight factor. At most, there was a 25 $\mu$m improvevent in the Standard Deviation.
See Figure \ref{fig:plane_residuals}
\begin{figure}[ht!]
  \hspace*{-1.5cm}                                                           
%  \vspace*{10.5cm}                                                           
  \centering
  \includegraphics[width=7.0in, height=4.0in]{residuals_compare.pdf}
  \caption{Plane residuals with the sigma-per-wire flag ON (\textcolor{red}{RED}) and OFF (\textcolor{navyblue}{BLUE})}
  \label{fig:plane_residuals}
\end{figure}
\newline
\noindent It is importnant to note that the sigma-per-wire flag and the parameter file are NOT included in \texttt{hallc$\_$replay}, since this is a very specific scenario that
ONLY applies to the Winter-2017/Spring-2018 Run Period. If users who took data during that time, and would like to try adding this flag and the parameter file,
the instructions are as follows:
\begin{itemize}
\item  Create the parameter file in path: \texttt{/PARAM/HMS/DC/hdc$\_$sigma$\_$per$\_$wire.param} \par
  \begin{minipage}{\linewidth}
    \centering
    \includegraphics[width=15cm]{param.png}
    \captionof{figure}{Snippet of sigma-per-wire parameter file.}
    \label{fig:param}
  \end{minipage}
\item  Include the parameter file in path: \texttt{/DBASE/HMS/general.param} \par
  \begin{minipage}{\linewidth}                                                          
    \centering
    \includegraphics[width=10cm]{gen_parm.png}
    \captionof{figure}{Snippet of general.param file.}
    \label{fig:gen_param}
  \end{minipage}
\end{itemize}
\noindent In case you need help or do NOT have the time to create this parameter file, I have uploaded both the script that produces the parameter file and the parameter file
itself to my github page. \\
\textbf{Link to script:} \\ \url{https://github.com/Yero1990/scientific_articles/blob/master/Spring_2018_Run/OutOfSync_TDC/write_sig_param.C} \\
\textbf{Link to parameter file:} \\\url{https://github.com/Yero1990/scientific_articles/blob/master/Spring_2018_Run/OutOfSync_TDC/hdc_sigma_per_wire.param}

\newpage
\onecolumn
\bibliography{template}
\bibliographystyle{acm}
